Data Analysis on Red Wine Quality by Ziyue Chi
========================================================

```{r global_options, include=FALSE}

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```



```{r include=FALSE, cache=FALSE, packages}

install.packages("ggplot2", dependencies = T, repos = 'https://cran.rstudio.com') 
install.packages("knitr", dependencies = T, repos = 'https://cran.rstudio.com')
install.packages("dplyr", dependencies = T, repos = 'https://cran.rstudio.com')
install.packages('GGally', repos = 'https://cran.rstudio.com')

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

```

```{r echo=FALSE, Load_the_Data}
# Load the Data

wine = read.csv('wineQualityReds.csv')

```


# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}

summary(wine)

```

This data set contains approximately 2,000 loans with 13 variables on each.


```{r echo=FALSE}

library(ggplot2)

library(gridExtra)

ggplot(aes(as.factor(quality)), data = wine) + geom_bar()

```

The level quality of wine in datasetis is normally distributed. The number of 
red wine with quality score 5 and 6 are 640 and 620 respectively.


```{r echo=FALSE}

grid.arrange(ggplot(wine, aes(x = 1, y = volatile.acidity)) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ),
             ggplot(wine, aes( x = volatile.acidity ) ) + 
                      geom_histogram(bins = 30 ), ncol = 2)


```

The level of volatile acidity is right skewed. The median is around 0.45

```{r echo=FALSE}

grid.arrange(ggplot(wine, aes(x = 1, y = alcohol)) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ),
             ggplot(wine, aes( x   = alcohol  ) ) + 
                   geom_histogram(bins=30 ),ncol=2)

```

The level of alcohol is fair distributed but slightly right skewed. The median 
is around 10.20.

```{r echo=FALSE}

grid.arrange(ggplot(wine, aes(x = 1, y = pH)) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ),
             ggplot(wine, aes( x   = pH  ) ) + 
                   geom_histogram(bins=30 ),ncol=2)

```

The scale of pH is normal distributed and most wines are between 3-4 on the
pH scale. The median is about 3.310.


```{r echo=FALSE}

grid.arrange(ggplot(wine, aes(x = 1, y = residual.sugar)) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ),
             ggplot(wine, aes( x   = residual.sugar  ) ) + 
                   geom_histogram(bins=30 ),ncol=2)

```

The residual sugar is strongly right skewed. It is important to mention that,
there is a large amount of outliers. The median is 2.200.

```{r echo=FALSE}

grid.arrange(ggplot(data = subset(wine, wine$residual.sugar < 3.66),
                    aes(x = 1, y = residual.sugar)) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ),
             ggplot(data = subset(wine, wine$residual.sugar < 3.66),
                    aes( x   = residual.sugar  ) ) + 
                   geom_histogram(bins=30 ),ncol=2)

```

After we calculate the upper fence which is 3.65, the refined plot provides a
better boxplot. However, the distribution plot raises some questions for me, 
why is there a gap at 2.7? Are there really no wine with that level of sugar?

```{r echo=FALSE}

grid.arrange(ggplot(wine, aes(x = 1, y = chlorides)) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ),
             ggplot(wine, aes( x   = chlorides  ) ) + 
                   geom_histogram(bins=30 ),ncol=2)

```

Similar to the last plot, there are a great amount of outliers in chlorides.
In addition, the distribution is strongly right skewed. The median is 0.07900.

```{r echo=FALSE}

grid.arrange(ggplot(subset(wine, (wine$chlorides < 0.13) &
                             (wine$chlorides > 0.04)),
                           aes(x = 1, y = chlorides)) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ),
             ggplot(wine, aes( x   = chlorides  ) ) + 
                   geom_histogram(bins=30 ),ncol=2)

```

After we calculate the upper fence and lower fence which is 0.13 and 0.04 
respectively, the refined plot provides a better boxplot. However, the
distribution plot raises some questions for me, why is there a gap at 2.7? 
Are there really no wine with that level of sugar?

# Univariate Analysis

### 1 What is the structure of your dataset?

There are about 1600 observations in the dataset with 11 features 
(fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, 
free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, alcohol).

-The quality of wines are normally distributed. 
-The distribution of most features in data is right skewed. 
-The maximum quality scale is 8.0 and the median scale is 6.0. 

### 2 What is/are the main feature(s) of interest in your dataset?

The main features I am interested in the data set are which chemical properties
influence the quality of red wines。 
I would like to determine which features would have a mainly affect on 
predicting the quality of wine.

### 3 What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Citric acid and alcohol. But I suspect that the level of volatile acidity 
plays a main role in affecting the quality of wine.

I also might provide a refined data that without outliers in chemical properties
such as sugar and chlorides.

### 4 Did you create any new variables from existing variables in the dataset?
Nope

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots}

wine_refine <- subset(wine, (wine$chlorides < 0.13) &
                             (wine$chlorides > 0.04) &
                        (wine$residual.sugar < 3.66))

library(GGally)
theme(axis.text = element_blank())

set.seed(888)
wine_subset <- wine[, c('volatile.acidity', 'alcohol', 'citric.acid',
                        'density', 'pH', 'residual.sugar',  'quality')]
names(wine_subset)
ggpairs(wine_subset, lower = list(continuous = wrap("smooth", method = "lm",
                                                    color = 'red')))
```

```{r echo=FALSE}

ggplot(data = wine_refine, aes(x=factor(quality), y=alcohol)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(color="red", fill="orange", alpha=0.5) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "blue", 
               shape = 8, 
               size = 4)

```

The correlation between level of alcohol and wine's quality is 0.476. The median
of the level of alcohol is increasing as the quality improved, from 9 to 12.
The graph indicates a positive relationship between level of alcohol and wine's
quality, especially for the wine with quality from level 5 to 6.

```{r echo=FALSE}

ggplot(data = wine_refine, aes(x=factor(quality), y=volatile.acidity)) +
  geom_jitter( alpha = 0.3 ) +
  geom_boxplot(color="red", fill="orange", alpha=0.5) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "blue", 
               shape = 8, 
               size = 4)

```

The correlation between level of alcohol and wine's quality is -0.4. The median
of volatile acidity is decreasing as the red wine's quality improved. It is from
0.8 down to about 0.4. The graph indicates a negative relationship between level
of volatile acidity and wine's quality.

```{r echo=FALSE}

ggplot(aes(alcohol, density), data = wine_refine) + geom_point(alpha = 0.2,
                                                        color = 'blue') +
  geom_smooth(method = 'lm', color = 'red')

```

The plot provides a negative trend (-0.5) between the level of alcohol and
density.

```{r echo=FALSE}

ggplot(aes(residual.sugar, density), data = wine_refine) +
  geom_point(alpha = 0.1, color = 'blue') +
  geom_smooth(method = 'lm', color = 'red')

```

The plot provides a positive trend (0.35) between the level of alcohol and
density.

# Bivariate Analysis

### 1 Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

From the plot above, alcohol has the strongest relationship with quality. 
While volatile acidity has a negative correlation with quality, but
the correlation is negative. 

### 2 Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Although density has a weak correlation with quality, the plot indicates a 
strong correlation with alcholo and residual sugar. Not surprisingly, there is 
a positive relationship between alcohol and density, and a negative relationship
between residual sugar and density.

### 3 What was the strongest relationship you found?

The negative relationship between volatile acidity and citric acid.


# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

ggplot(aes(x = alcohol, 
            y = volatile.acidity, color = factor(quality)), 
        data = wine_refine) +
       geom_point(alpha = 0.8, size = 1) +
       geom_smooth(method = "lm", se = FALSE,size=1)  +
   scale_color_brewer(type='seq',
                    guide=guide_legend(title='Quality')) + theme_dark()
```

Confront with chemical properties like alcohol and volatile acidity, the 
influence of these chemical properties on quality of red wine is not significant.
The regression lines in plot above show that only red wine with quality 3 and 8
are evident. It is important to mention that, there are few number of red wine
with quality 3 and 8, compare with red wine with quality 5 and 6.


```{r echo=FALSE}

ggplot(aes(x = density, 
           y = alcohol, color = factor(quality)), 
       data = wine_refine) +
      geom_point(alpha = 0.8, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality')) + theme_dark()
  
```

Conversely, given the plot above, the influence of alcohol on quality is 
significant, given the constant level of surgar. It is obvious that the 
influence is significant to all quality levels.

```{r echo=FALSE}

ggplot(aes(x = density, 
           y = volatile.acidity, color = factor(quality)), 
       data = wine_refine) +
      geom_point(alpha = 0.8, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality')) + theme_dark()

```

While, if we control the level of surgar, the influence of volatile acidity
on quality is not as significant as the alcohol.

# Multivariate Analysis

### 1 Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

While if I combine the two chemical properties like alcohol and volatile acidity
their influence on level of red wine quality is weakened. It seems the 
influence only works form extremly good wien and extremly bad wine.


### 2 Were there any interesting or surprising interactions between features?

If I control the level of surgar, the influence of alcohol and volatile acidity
on quality is as same as last section suggested. While if I put alcohol and 
volatile acidity together, the influence is not as significant as last section
showed.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

ggplot(wine, aes(x = 1, y = residual.sugar)) + geom_jitter(alpha = 0.1 ) + 
               ylab('Residual Sugar (Gram/Liter)') +
               ggtitle("The Boxplot for Residual Sugar ") + 
               geom_boxplot(alpha = 0.2, color = 'red' )

```

### Description One

The plot shows a there are a great amount outliers of sugar. However, it turns 
out that there are not important to the influence to quality, especially when
confront with other chemical property such as alcohol.

### Plot Two
```{r echo=FALSE, Plot_Two}

ggplot(aes(x = density, 
           y = alcohol, color = factor(quality)), 
       data = wine_refine) +
      geom_point(alpha = 0.8, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality')) + 
  ylab("Alcohol") +
  xlab("Density") +
  ggtitle("How Alcohol and Density will influence Quality ") + theme_dark()
```

### Description Two

The plot shows a negative influence of volatile acidity on quality. Similarly to
last plot, it seems like the great amount of outliers does not affect alcohol's 
influence.

### Plot Three
```{r echo=FALSE, Plot_Three}

ggplot(aes(x = alcohol, 
           y = volatile.acidity, color = factor(quality)), 
       data = wine_refine) +
      geom_point(alpha = 0.8, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality')) +
  ylab("Volatile Acidity (g/dm^3)") +
  xlab("Alcohol") +
  ggtitle("Combine with Alcohol and Volatile Acidity, How the influence on 
          Quality might looks like ") + theme_dark()

```

### Description Three

I would like to see how the a large amount of outliers will affect the influence.
From the graph above, with Alcohol and Volatile Acidity, it is not clear that 
how those two chemical properties will influence the red wine that scored 
from 4 to 7.

------

# Reflection

This data contains approximately 2,000 loans with 13 variables on each. My main
question is which chemical property would influence the quality of red wine, and
how actually the correlation is. 

My analysis suggests that both alcohol and Volatile Acidity play an important
role in influencing the quality, but in different direction. The influence of 
alcohol is strong if the level of surgar is controlled, while this argument
does not hold for volatile acidity. Inaddition, we need to be aware of the 
large amount of outliers in chemical properties such lik sugar, alcohol and 
volatile acidity, especially for red wine scored 4, 5 and 6. A multivariate plot
shows that confront with Volatile Acidity and alcohol, the influence is only 
significant for with quality 3 and 8. 

In order to do further research, the data need to provide more observations, 
and a more advanced regression is desired.

The struggle of my analysis is that the correlation between the given chemical
properties and quality of red wine. It looks like easy, single regression might
deal with that, but it becomes more complicate than I originally thought when I 
combine more chemical properties. It also looks trick if I try to apply
multivariable regression, since several of contral variables are linearly relate 
to others.

Data visulization makes my life easier, based on some plots, I figure out, there
are plenty outliers among those chemical properties. So I wonder if this result
might shed some lights on my analysis. The result show that the influence of 
alcohol is robust especially when they confront with sugar, while volatile 
acidity is not as robust as alcohol is. In addition, they contribute to the wine
with great quality. So I am wondering that if their combined chemical reaction
is a threshold for making great wine. Secondly, I can, of course, get rid of the
outliers. But, due to the limitation of number of observation in dataset. And 
this raises a question that there is a gap in the red wine when I control the 
number of outliers, I am not able to deal with it at the moment.I would suggest
that more and more observations is desired.





